---
title: "README"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This package is designed to help with common tasks in the Social Security Scotland statistics team. 

# Getting started
## Using the package
To install the package directly from GitHUb, open RStudio and run:

```{r install, message = FALSE}
devtools::install_github("ScotGovAnalysis/sssstats",
               upgrade = "never",
               build_vignettes = TRUE)
```

For details of the functions available, run `??sssstats` and click `Index` at the bottom of the help page to access the package documentation.

## Contributing to the package
Contributions to the package are very welcome. If you would like to contribute, please fork the directory and open a pull request.

Details of how to do so are available [here](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/proposing-changes-to-your-work-with-pull-requests/creating-a-pull-request).
